Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-03-29 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs: more editing

We start by presenting our results for EdGB gravity.
%
In Fig.~\ref{fig:edgb_exec_sum} we show the margelized posteriors of the coupling constant $\ell_{\rm GB}$,
for GW150914 (top panel) and GW200129 (middle panel), with and without the spin correction to the $(2,2,0)$
QNM. The bottom panel panel shows the joint posterior combining both events.
%
We see for both events events the $N_{\rm max} = 0$ posteriors are characterized by a peak away from zero.
%
This does not mean that we are observing a deviantion from GR. We recall that the deviations from GR in the ParSpec framework
are controlled by $\gamma$, which here reads,
%
\begin{equation}
    \gamma_{\rm GB} = \left(\frac{c^2 \ell_{\rm GB}}{ G M_{\rm s}}\right)^4\,.
\end{equation}
%
As shown in Fig.~xxx, $\gamma_{\rm GB}$ does indeed have a posterior with maximum support near zero, indicating consistency
of our data with GR.
%
We also observe that the inclusion of the spin corrections (i.e., $N_{\rm max} = 1$) displaces the posteriors towards
smaller values of $\ell_{\rm GB}$.

As we stressed in Sec.~\ref{sec:remarks}, before we can draw any conclusions on
the allowed values for $\ell_{\rm GB}$ from our parameter estimation study, we
must first check whether the ``EFT''~\eqref{eq:eft_bound} and
``ParSpec''~\eqref{eq:parspec_bound} bounds are satisfied.
%
We do the former in Fig.~\ref{fig:edgb_cdf}.

>>>>

We start by presenting our results for EdGB gravity.
%
In Fig.~\ref{fig:edgb_exec_sum} we show the margelized posteriors of the coupling constant $\ell_{\rm GB}$,
for GW150914 (top panel) and GW200129 (middle panel), with and without the spin correction to the $(2,2,0)$
QNM. The bottom panel panel shows the joint posterior combining both events.
%
We see for both events events the $N_{\rm max} = 0$ posteriors are characterized by a peak away from zero.
%
This does not mean that we are observing a deviantion from GR. We recall that the deviations from GR in the ParSpec framework
are controlled by $\gamma$, which here reads,
%
\begin{equation}
    \gamma_{\rm GB} = \left(\frac{c^2 \ell_{\rm GB}}{ G M_{\rm s}}\right)^4\,.
\end{equation}
%
As shown in Fig.~xxx, $\gamma_{\rm GB}$ does indeed have a posterior with maximum support near zero, indicating consistency
between data and GR.
%
We also observe that the inclusion of the spin corrections (i.e., $N_{\rm max} = 1$) displaces the posteriors towards
smaller values of $\ell_{\rm GB}$.

As we stressed in Sec.~\ref{sec:remarks}, before we can draw any conclusions on
the allowed values for $\ell_{\rm GB}$ from our parameter estimation study, we
must first check whether the ``EFT''~\eqref{eq:eft_bound} and
``ParSpec''~\eqref{eq:parspec_bound} bounds are satisfied.
%
We verify the former in Fig.~\ref{fig:edgb_cdf}. In top (bottom) panel we shown the cumulative density function
of the $\ell_{\rm GB}$ posteriors, obtained evaluating the integral~\eqref{eq:def_cdf} with $\Lambda_{\rm EFT}(\varepsilon, \mathfrak{m})$,
with the mass scale set by the secondary's mass (i.e., $\mathfrak{m} = m_2$, dashed lines) or the remnant's mass ($\mathfrak{m} = M_f$, solid lines), with varying $\epsilon$.

### 2022-03-29 papers/paper_alt_theor_bounds/sec_pe.tex
hs: more editing

First, as we explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be considered as an effective field theory, meaning it should be
considered valid only below an energy scale, or equivalently, a length scale.
%
As a cut-off length scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon, \mathfrak{m}) = \varepsilon \, G \mathfrak{m} / c^2 \,,
\label{eq:def_cutoff}
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the mass scales involved in the problem.
%
We also momentarily restored factors of $c$ and $G$ to emphasize that $\Lambda_{\rm EFT}$ has
dimensions of length and hence can be compared to the coupling strength $\ell$.
%
In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, $\varepsilon = 1/2$ was used,
but here we consider $\varepsilon \in [0, 1]$.
%
We will say that a bound has been placed on $\ell$, if most of the support of
the posterior distribution function $P(\ell)$ is in the interval
$[0, \, \Lambda_{\rm EFT}(\varepsilon, \mathfrak{m})]$.
%
In practice, this can be quantified through the cumulative distribution function
(CDF) associated to the marginalized posterior distribution $P(\ell)$,
%
\begin{equation}
P(\ell \leqslant \ell_{\rm max}) = \int_{-\infty}^{\,\ell_{\rm max}} P(\ell') \, \dd \ell'\,.
\label{eq:def_cdf}
\end{equation}
%
For instance, we will demand that for a bound with 90\% credibility to be meaningfully placed that,
%
\begin{equation}
% \textrm{CDF}(\ell) \leqslant \textrm{CDF}(T_{\rm EFT}(\varepsilon; \, \mathfrak{m})).
P(\ell \leqslant \Lambda_{\rm EFT}) \leqslant 0.9 \,,
\quad \textrm{(EFT bound)},
\label{eq:eft_bound}
\end{equation}
%
and likewise for other percentiles.

>>>>

First, as we explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be considered as an effective field theory, meaning it should be
considered valid only below an energy scale, or equivalently, a length scale.
%
As a cut-off length scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon, \mathfrak{m}) = \varepsilon \, G \mathfrak{m} / c^2 \,,
\label{eq:def_cutoff}
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the mass scales involved in the problem.
%
We also momentarily restored factors of $c$ and $G$ to emphasize that $\Lambda_{\rm EFT}$ has
dimensions of length and hence can be compared to the coupling strength $\ell$.
%
In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, $\varepsilon = 1/2$ was used,
but here we consider $\varepsilon \in [0, 1]$.
%
We will say that a bound has been placed on $\ell$, if most of the support of
the posterior distribution function $P(\ell)$ is in the interval
$[0, \, \Lambda_{\rm EFT}(\varepsilon, \mathfrak{m})]$.
%
In practice, this can be quantified through the cumulative distribution function
(CDF) associated with the marginalized posterior distribution $P(\ell)$,
%
\begin{equation}
P(\ell \leqslant \ell_{\rm max}) = \int_{-\infty}^{\,\ell_{\rm max}} P(\ell') \, \dd \ell'\,.
\label{eq:def_cdf}
\end{equation}
%
For instance, we will demand that for a bound with 90\% credibility to be meaningfully placed that
up to $\ell_{\rm max} = \Lambda_{\rm EFT}$ we have,
%
\begin{equation}
% \textrm{CDF}(\ell) \leqslant \textrm{CDF}(T_{\rm EFT}(\varepsilon; \, \mathfrak{m})).
P(\ell \leqslant \Lambda_{\rm EFT}) \leqslant 0.9 \,,
\quad \textrm{(EFT bound)},
\label{eq:eft_bound}
\end{equation}
%
and likewise for other credibility percentiles.
